---
title: "An R package for SPRE"
author: ""
output: rmarkdown::github_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we show two examples of SPRE analysis. The first example can be found in Weissman-Miller, D., Miller, R. J., & Shotwell, M. P. (2017), while the second example can be found in Weissman-Miller, D. (2013). The references are given as follow.

Weissman-Miller, D. (2013). Novel point estimation from a Semiparametric Ratio Estimator (SPRE): Long-term health outcomes from short-term linear data, with application to weight loss in obesity. International Journal of Biostatistics, 9(2): 175-184 DOI: http://dx.doi.org/10.1515/ijb-2012-0049.

Weissman-Miller, D., Miller, R. J., & Shotwell, M. P. (2017). Translational Research for Occupational Therapy: Using SPRE in Hippotherapy for Children with Developmental Disabilities. Occupational therapy international, 2017.

## First example

### Load package

```{r}
library(SPRE)
```

### Load HEAT data

```{r}
data("HEAT_stat")
```

### Step 1: Determining the most significant F statistic by stepwise regression

```{r}
skip_indices<-1:2
step_reg<-stepwise_regression(HEAT_stat$Session,HEAT_stat$FData,skip_indices)
```

### Step 2: Evaluating the change point and estimating parameter

Here, the change point is identified as the 11th index.

```{r}
SPRE_mod<-SPRE(step_reg,chptidx=11,lastidx=14)
```

### Step 3: Estimating/predicting long-term outcomes

We will need to supply the following things to the function:

- Times/sessions the predictions will be evaluated at: xpred

- Shape parameter: kshape

- Change point: tau

- Prior response: theta1

```{r}
kshape<-SPRE_mod$kshape
idxmaxFstat<-which.max(SPRE_mod$step_reg$fDist[,"Fstat"])
tau<-SPRE_mod$step_reg$fDist[idxmaxFstat,"x"]
theta1<-HEAT_stat$FData[SPRE_mod$chptidx]
xpred<-7:46
SPRE_pred<-predict_SPRE(xpred=xpred,kshape=kshape,tau=tau,
                        theta1=theta1,predtype="rolling")
```

### Additional analyses and checks

#### Residuals plot

```{r,fig.width=7,fig.height=7}
plot_residuals(SPRE_mod)
```

#### Plot probability distribution
```{r,fig.width=7,fig.height=7}
plot_probability(SPRE_mod,SPRE_pred,xmin=2)
```

#### Plot things to check weibull distribution
```{r,fig.width=7,fig.height=7}
plot_weibull(SPRE_mod,SPRE_pred)
```

#### Check prediction stability
```{r,fig.width=7,fig.height=7}
prediction_stability(SPRE_pred)
```

## Second example

### Load package

```{r}
library(SPRE)
```

### Input data

```{r}
x<-c(0,0.25,0.5,1.0,1.5,1.75,2.0,2.5,3.0,3.5,4.0,4.5,5.0,5.5,6.0,6.5,7.0,7.5,8.0,9.0,10.0)
y<-c(0,0.5,1.0,2.5,3.5,4.5,6.0,6.5,7.5,8.0,9.0,9.5,
     10.0,10.5,11.0,10.8,10.75,10.6,10.5,10.4,10.25)
skip_indices<-c(1:6,12,14,16,18,20)
```

### Step 1: Determining the most significant F statistic by stepwise regression

```{r}
step_reg<-stepwise_regression(x,y,skip_indices)
```

### Step 2: Evaluating the change point and estimating parameter

The change point is identified as the 11th index.

```{r}
SPRE_mod<-SPRE(step_reg,chptidx=11)
```

### Step 3: Estimating/predicting long-term outcomes

```{r}
kshape<-SPRE_mod$kshape
idxmaxFstat<-which.max(SPRE_mod$step_reg$fDist[,"Fstat"])
tau<-SPRE_mod$step_reg$fDist[idxmaxFstat,"x"]
theta1<-y[SPRE_mod$chptidx]
xpred<-c(4,4.5,5,5.5,6,7,8,9,10)
SPRE_pred<-predict_SPRE(xpred=xpred,
                        kshape=kshape,tau=tau,theta1=theta1)
```

### Additional analyses and checks

#### Residuals plot

```{r,fig.width=7,fig.height=7}
plot_residuals(SPRE_mod)
```

#### Plot probability distribution
```{r,fig.width=7,fig.height=7}
plot_probability(SPRE_mod,SPRE_pred,xmin=2)
```

#### Plot things to check weibull distribution
```{r,fig.width=7,fig.height=7}
plot_weibull(SPRE_mod,SPRE_pred)
```

#### Check prediction stability
```{r,fig.width=7,fig.height=7}
prediction_stability(SPRE_pred,npred=length(xpred))
```
